Diagnosing Simple Wind Models of Cool Gas Outflows as Viewed by Metal-Line Absorption and Emission

### 2010-10-11 Winds/papers/Theory/First/Submit/ms.tex
blah
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5263 ac11df5e-a788-c23b-a8e1-c758c323924d

\title{Diagnosing Simple Wind Models of Cool Gas Outflows through
  Metal-Line Absorption and Emission}

@@

\begin{abstract}
We analyze the absorption and emission-line profiles produced by a set
of simple, cool gas wind models motivated by galactic-scale outflow
observations.  We focus on the \mgiid\ doublet and \ion{Fe}{2} UV1
multiplet at $\lambda \approx 2600$\AA, but the results are applicable
to most other transitions used to study outflows (e.g.\ \ion{Na}{1},
\ion{H}{1} \lya, \ion{Si}{2}).   We implement monte carlo radiative
transfer techniques to properly track the propogation of scattered and
flourescent photons to [produce] 1D spectra and 2D spectral images.
[Should we make the codes publically available?  Upon request?]
[Mention fiducial model here?]
By design, the \ion{Mg}{2} doublet exhibits strong, blue-shifted
absorption but also comparably strong line-emission at roughly
systemic velocity.  This line-emission `fills-in' the absorption at
velocieties near systemic, significantly reducing the inferred
opacity. Analysis of cool gas outflows that entirely ignores this
line-emission may incorrectly conclude that the gas is partially
covered, has a significantly lower peak optical depth, and/or that a
gas with velocities near systemic (e.g.\ from the ISM or even
an infalling component) is absent.   
The \ion{Fe}{2} resonance lines show similar behaviour, however 
these lines are connected to fine-structure levels that are optically
thin to the medium.  As a result, the resonant line-emission is
reduced (at the `expense' of significant fine-structure emission) and
the absorption lines more faithfully reproduce the intrinsic opacity
of the wind.   [Together these results naturally explain the
absorption and emission-line characteristics observed for star-forming
galaies at $z<1$.]  
We examine outflows with a range of power-law density and velocity
profiles and a radiative-driven wind model.  The resultant \ion{Mg}{2}
and \ion{Fe}{2} profiles show a diverse complexity of absorption and
emission characteristics but overall have qualitatively similar
features.  We also study a scenario purported to describe the outflows
of $z \sim 3$ Lyman break galaxies and argue this model fails to
reproduce the observations when (previously ignored0 scattered photons
are properly considered.   Strong line-emission from metal-line
transitions is a generic prediction of wind models, even in the
presence of dust and anisotropic configurations.  We calculate the
surface brightness profiles of this emission which motivates
spatially-extended spectroscopic observations with current and
next-generation integral-field-unit instruments.

@@

Indeed, line-emission from low-ion transitions has been reported from 
star-forming galaxies that exhibit cool gas outflows.  
This includes emission related to \ion{Na}{1} \citep{phillips93,cth+10},
\ion{Mg}{2} emission \citep{wcp+09,rwk+10}, 
\ion{Si}{2}* emission from LBG galaxies \citep{shapley03},
and, most recently, significant \feiis\ emission \citep{rubin+10c}. 
A variety of origins have been proposed for this
line-emission including AGN activity, recombination in \ion{H}{2}
regions, and back-scattering off the galactic-scale wind.
Our results suggest that the majority of the observed line-emission is 
from scattered photons in the cool gas outflows of these star-forming
galaxies.  Indeed, line-emission should be generated by all 
galaxies driving a wind.

On the other hand,
there are many examples of galaxies where blue-shifted absorption is
detected yet the authors report no significant line emission.  This
includes \ion{Na}{1} flows \citep{rvs05a,martin05,smn+09}, 
\ion{Mg}{2} and \ion{Fe}{2} absorption in
ULIRGS \citep{mb09}, and the extreme \ion{Mg}{2} outflows identified
by \cite{tmd07}.  Similarly, there have been no reports of
resonance line-emission from low-ion transitions in the
LBGs\footnote{\citet{psa+00}
  reported and analyzed \ion{C}{4} line-emission from the lensed LBG
  cb58, attributing this flux to stellar winds.  We caution, however,
  that the galactic-scale wind proposed to exist for cb58 should also
  have contributed to the \ion{C}{4} line-emission.}, and many
$z \lesssim 1$ galaxies show no detectable \ion{Mg}{2} or \ion{Fe}{2}*
emission despite significant blue-shifted absorption (Rubin et al., in prep). 
These non-detections appear to contradict a primary conclusion of 
this paper.   We are motivated, therefore, to
reassess several of the effects that can reduce the line-emission
and consider whether these prevent its detection for many
star-forming galaxies.

Dust is frequently invoked to explain the suppression of line-emission
for resonantly trapped transitions (e.g.\ \lya).  Indeed, a photon
that is trapped for many scatterings within a dusty medium will 
be preferentially extincted relative to a non-resonant photon.  In
$\S$~\ref{sec:dust}, we examined the effects of dust extinction on 
\ion{Mg}{2} and \ion{Fe}{2} emission.  The general result
(Figure~\ref{fig:dust_tau}) is a modest
reduction in flux that scales as (1+\taud)$^{-1}$ instead of
exp(--\taud). Although the \ion{Mg}{2} photons are resonantly
trapped, they require only one to a few scatterings
to escape the wind which limits the effects of
dust.  This reflects the moderate opacity of the \ion{Mg}{2}
doublet (e.g.\ relative to \lya) and also the velocity law of the
fiducial wind model.
As we have discussed, in scenarios where the \ion{Mg}{2} photons are
more effectively trapped dust further suppresses
the emission (e.g.\ the ISM+wind model from $\S$~\ref{sec:ISM}).
In contrast to \ion{Mg}{2}, 
the \ion{Fe}{2} resonance photons may be
converted to non-resonant photons that freely escape the wind.
Therefore, a wind model that traps \ion{Mg}{2}
photons for many scatterings does not similarly trap 
\feiid\ photons.  The effects of dust are much 
reduced and, by inference, the same holds for any other set of
transitions that are coupled to a fine-structure level (e.g.\
\ion{Si}{2}/\ion{Si}{2}*).  In summary,
dust does reduce the line flux relative to the continuum, but 
it generally has only a modest effect on the predicted \ion{Mg}{2}
emission and a minor effect on \feiis\ emission.

Another factor that may reduce line-emission is a
non-isotropic wind model. The flux is lower, for example, if one
eliminates the backside to the wind (e.g.\ the source itself could
shadow a significant fraction of the backside).  
For our fiducial model, many of the photons scattered into our
sightline arise nearly normal to the sightline such that a biconicial
wind has significantly lower line emission
($\S$~\ref{sec:anisotropic}). 
For the anisotropic winds explored in this paper,
the line-emission scales 
roughly as $\Omega/4\pi$ where $\Omega$ is the angular extent of the
wind\footnote{There is also a first order dependence on the actual
  orientation of the wind, with maximal emission when the backside is
  fully viewed.}.  Because we require that the wind 
points toward us, it is difficult to reduce $\Omega$ much
below $2 \pi$ and, therefore, anisotropy reduces the
emission by a factor of order unity.  
Similarly, an anisotropically emitting source would modify the predicted
line-emission.  For example, if the backside of the galaxy were
brighter/fainter then would one predict brighter/fainter emission
relative to the observed continuum.
To reduce the emission, the brightest regions
of the galaxy would need to be oriented toward Earth.  Although this
would not be a
generic orientation, most spectroscopic samples are magnitude-limited
and biased towards detecting galaxies when observed from their
brightest viewpoint.  In principle, this might imply a further
reduction of order unity.
{\bf [Put comments on biconical here]}

Together, dust and anisotropic models may reduce the line-emission
of \ion{Mg}{2} and \ion{Fe}{2}
by one to a few factors of order unity.  For some galaxies, these
effects may help to explain the absence of significant line-emission,
but these would generally
not be sufficient to preclude the detection.

There is another (more subtle) effect that could
greatly reduce the observed line-emission: slit-loss.  As
described in Figure~\ref{fig:fiducial_ifu_mgii}, 
emission from the wind is spatially extended with
a non-zero surface brightness predicted at large radii. This implies 
a non-negligible luminosity emitted well beyond the galaxy.  
The majority of observations of star-forming galaxies to date have
been taken through a spectroscopic slit designed to cover
the brightest continuum regions of the galaxy. Slits with angular
extents of $1-1.5''$ subtend roughly $8-15$\,kpc for galaxies at $z>0.5$.
Therefore, a $1''$ slit covering a galaxy with our fiducial wind would cover
less than half the full wind.  The result is reduced line-emission
when compared to the galaxy continuum.

We explore the effects of slit-loss as follows.
We model the slit as a perfect, infinitely-long rectangle centered on the
galaxy.  We then tabulate the equivalent width of the line-emission
through slits with a range of widths, parameterized by \rtt: twice the radius
where the wind has a Sobolev optical depth of $\tau_S = 0.2$. 
The line-emission will be weak beyond
this radius because the photons have only a low probability of
being scattered.  For the fiducial wind model, 
the \mgiia\ transition has $\mrtt \approx 15$\,kpc 
(Figure~\ref{fig:fiducial_nvt}).  
The predicted equivalent width in emission \ewe\ relative to the absorption
equivalent width \ewabs\ is presented in
Figure~\ref{fig:obs_slit} for a series of transitions for the fiducial
wind model (black curves).
[Add radiative model]
The \ewe/\ewabs\ curves rise very steeply with increasing slit width and then
plateau when the slit width reaches $\approx \mrtt$.  
For the fiducial model, the emission is concentrated toward the source; this
derives from the density and velocity profiles but is also a simple
consequence of geometric projection.
Figure~\ref{fig:obs_slit} reveals similar results for other wind
models.
All of the curves rise steeply enough that 
the effects of slit-loss are minor (order unity)
unless the slit-width is very small.   
Nevertheless, the results do motivate
extended aperture observations, e.g.\ integral field unit (IFU)
instrumentation that would map the wind both spatially and spectrally.

@@

Figure~\ref{fig:obs_ew} demonstrates one observational
consequence: reduced measurements for the absorption
equivalent width \ewabs\ of the flow.  In the case of \ion{Mg}{2},
which has the most affected transitions, \ewabs\ is reduced by
$30-50\%$ from the intrinsic equivalent width.  In turn, one may derive 
a systematically lower optical depth or velocity extent for the wind,
and therefore a lower total mass and kinetic energy.  
The effects are most pronounced for wind
scenarios where the peak optical depth occurs near $\mdv = 0 \mkms$.
Geometric projection limits the majority of scattered photon emission to
have $|\mdv| < 200 \mkms$; therefore,  the absorption profiles
are filled-in primarily at these velocities.

@@

These effects for \ion{Mg}{2} absorption are reduced
for the \ion{Fe}{2} absorption profiles, especially for \feiia.  This
is because a significant fraction (even a majority) of the
\ion{Fe}{2} emission is florescent \ion{Fe}{2}* emission at longer
wavelengths which does not affect the absorption profiles. 
Therefore, the \ion{Fe}{2} absorption equivalent widths (\ewabs) more
closely follow the intrinsic values, one derives more accurate peak
optical depths, and the absorption kinematics more faithfully reflect
the motions of the flow.  Regarding the last point, one also
predicts a velocity offset between the \ion{Fe}{2} and \ion{Mg}{2}
absorption-line centroids (Figure~\ref{fig:obs_vtau}).
%because scattered photons
%preferentially fill-in
%absorption at $v \sim 0 \mkms$ for the \ion{Mg}{2} transitions.  
This affects the analysis of gas related to the ISM of the galaxy and
also material infalling at a modest speed.  
Figure~\ref{fig:obs_vtau} also emphasizes that
the \ion{Mg}{2} profiles may misrepresent the kinematics of the bulk of
the gas.  Analysis of these lines, without consideration of
line-emission, would lead to incorrect conclusions on the energetics
and mass flux of the wind.

@@

The most obvious measures provided by line-emission analysis are
estimates
of the size and morphology of the outflow \citep{rubin+10c}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths.  The principle
challenge is to achieve sufficient sensitivity to detect the predicted
low
surface brightness emission. As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds that at the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. 
For the fiducial wind model, for example, the azimuthally integrated
flux falls by only a factor of 10 from $r = 1$ to 15\,kpc.
Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for this analysis because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  The \ion{Mg}{2} lines (especially \mgiib) frequently has the
highest peak and integrated fluxes (Figure~\ref{fig:obs_sb}).  
On the other hand, the \ion{Mg}{2} transitions are more susceptible to
the effects of dust extinction and for some galaxies \feiis\ emission
could be dominant.  In either case, the study of spatially-extended
line-emission from photons scattred by the galactic-scale wind offers
an direct means to study the morphology and radial extent of these
phenomena.  The next generaly of large-format optical and infrared
integral-field-units are well suited to this scientific endeavor
(e.g.\ KCWI on Keck, MUSE on the VLT).

The kinematic characteristics of the line-emission also offer insight into physical
characteristics of the wind.
In $\S$~\ref{sec:dust}, we emphasized that dust extinction
preferentially suppresses photons scattered off the backside of the
wind (e.g.\ $\mdv > 0 \mkms$) so that the line-centroid is shifted to
negative velocities.  Similarly, anisotropic models with reduced
emission off the backside yield emission lines that are centered
blueward of the galaxy's systemic redshift.  These effects are most
prominent in the \feiis\ emission.  The centroids of the \ion{Mg}{2}
lines, meanwhile, may be modified by absorption, e.g.\ a wind with
speeds exceeding 600\kms\ will have \mgiib\ scatter \mgiia\ photons.
Therefore, the line centroids are sensitive to the optical depth
profile of the wind.  Although these effects are modest (tens of
\kms), they can be [recorded] by modern resolution spectroscopy.

We also find that the emission-line widths reveal
characteristics of the wind.  In particular, the \feiis\ emission-line 
widths are
sensitive to the optical depth of the wind at large velocities from
systemic.  This is illustrated in Figure~\ref{fig:obs_edelv} which plots
the 90\%\ velocity width $\Delta v_e$ of \feiic\ versus the velocity
where the wind optical depth is maximal ($v_\tau$).  It is evident
that $\Delta v_e$ increases with $v_\tau$; a large measured width for
\feiis\ emission requires a wind profile with large optical depths at
$\mdv < -200 \mkms$.  The broadening of \feiis\ emission occurs
because this emission is dominated by single scatterings 
which trace all components of the wind.  {\bf [Also the optical depth profile of
  FeII]}

Lastly, the line flux ratios of pairs of transitions 
are sensitive to characteristics of the wind.
The most obvious example is the relative emisison of the \mgiid\
doublet.  Specifically, an outflow whose velocity exceeds the doublet
spacing ($\Delta v \approx 770 \mkms$) will convert \mgiia\ photons
into \mgiib\ emission resulting in a flux ratio that is inverted
relative to the intrinsic optical depth profiles (e.g.\
Figure~\ref{fig:norm}).  Another important example is the relative
flux of \feiis\ emission relative to \ion{Mg}{2}.  
A dusty medium with velocity near systemic (e.g.\ an
ISM component) may significantly suppress \ion{Mg}{2} emission yet
have a significant \feiis\ line flux because the latter is not
resonantly trapped.  [One last example?]

@@

In this paper, we have explored the predicted absorption and
emission-line profiles of a set of simple galactic-scale outflow
models.  Our analysis focused on the \mgiid\ and \ion{Fe}{2} UV1
multiple of rest-frame UV transitions, but the results translate to
most other lines thus far used to probe these phenomna.  Our primary
findings are summarized as follows:

@@

\item Isotropic, dust-free wind models conserve photon flux.
  Therefore, the blue-shifted absorption-line profiles commonly
  observed in star-forming galaxies are predicted to be accompanied by
  emission-lines with similar equivalent width.

\item The line-emission occurs preferentially at the systemic velocity
  of the galaxy and `fills-in' the absorption profiles at velocity
  offsets $|\mdv| < 200 \mkms$ from systemic.  For transitions that
  are only coupled to the ground-state (e.g.\ \ion{Mg}{2}, \lya,
  \ion{Na}{1}), one predicts this emission to significantly fill-in
  the absorption near systemic.

@@

\item Resonance transitions that are strongly coupled to
non-resonant lines (e.g.\ \ion{Fe}{2}, \ion{Si}{2}) offer the best
characterization of the opacity of the wind and also any component
with velocity near systemic.  The emission will be domianted by the
optically-thin, fine-structure transitions.

\item  Dust extinction has only modest effects on the profiles
  in models where one to only a few scatterings are predicted.  Models
  with high opacity at small radii and at systemic velocity (e.g.\ an
  optically thick, ISM component)  can effecitvely extinguish
  resonantly trapped emission (\ion{Mg}{2}) but have more modest
  effect on non-resonant lines (\feiis).

\item We examined two scenarios designed to mimic the wind model
  proposed by S10 for $z \sim 3$ Lyman break galaxies.  These models
  predict substantial line-emission from scattered photons that
  greatly modify the predicted line-profiles.  As such, 
  these models do not reproduce the observed line-profiles.

\item In fact, the
\feiia\ profile very nearly matches the profile without re-emission 
(compare to the dotted lines); this transition provides a
very good description of the intrinsic ISM+wind optical depth profile.  
We conclude that resonant transitions that are coupled to (multiple)
non-resonant, electric dipole transitions offer the best
diagnosis of ISM absorption.

>>>>

\title{Diagnosing Simple Wind Models of Cool Gas Outflows as Viewed by
  Metal-Line Absorption and Emission}

@@

\begin{abstract}
We analyze the absorption and emission-line profiles produced by a set
of simple, cool gas wind models motivated by galactic-scale outflow
observations.  
We examine outflows with a range of power-law density and velocity
profiles and a radiative-driven wind model.  
We focus on the \mgiid\ doublet and \ion{Fe}{2} UV1
multiplet at $\lambda \approx 2600$\AA, but the results are applicable
to other transitions that trace outflows (e.g.\ \ion{Na}{1},
\ion{H}{1} \lya, \ion{Si}{2}).   We implement monte carlo radiative
transfer techniques that track the propogation of scattered and
flourescent photons to generate 1D spectra and 2D spectral images.
[Should we make the codes publically available?  Upon request?]
[Mention fiducial model here?]
As expected, the \ion{Mg}{2} doublet exhibits strong, blue-shifted
absorption but one also predicts strong line-emission at roughly
the systemic velocity.  This line-emission `fills-in' the absorption, 
significantly reducing the effective
opacity near systemic. Analysis of cool gas outflows that entirely ignores this
line-emission may incorrectly infer that the gas is partially
covered, measure a significantly lower peak optical depth, and/or
conclude that gas at systemic velocity is absent   
(e.g.\ an ISM or slowly infalling component). 
The \ion{Fe}{2} resonance lines show similar behaviour, however 
these lines are connected to optically thin transitions to
fine-structure levels. As such, the \ion{Fe}{2} 
absorption lines more faithfully reproduce the intrinsic opacity
of the wind.   Together these results naturally explain the
absorption and emission-line characteristics observed for star-forming
galaies at $z<1$.
We also study a scenario purported to describe the outflows
of $z \sim 3$ Lyman break galaxies and argue this model fails to
reproduce the observations when (previously ignored) scattered photons
are properly considered.   
%Strong line-emission from metal-line
%transitions is a generic prediction of wind models, even in the
%presence of dust and anisotropic configurations.  
The surface brightness profiles of emission from metal-line
transitions offers a unique means of assessing the morphology and size
of galactic-scale winds motivating deep,
spatially-extended spectroscopic observations.

@@

Indeed, line-emission from low-ion transitions has been reported from 
star-forming galaxies that exhibit cool gas outflows.  
This includes emission related to \ion{Na}{1} \citep{phillips93,cth+10},
\ion{Mg}{2} emission \citep{wcp+09,rwk+10}, 
\ion{Si}{2}* emission from LBG galaxies \citep{shapley03},
and, most recently, significant \feiis\ emission \citep{rubin+10c}. 
A variety of origins have been proposed for this
line-emission including AGN activity, recombination in \ion{H}{2}
regions, and back-scattering off the galactic-scale wind.
Our results suggest that the majority of the observed line-emission is 
from scattered photons in the cool gas outflows of these star-forming
galaxies.  Indeed, line-emission should be generated by all 
galaxies driving a cool gas outflow.

On the other hand,
there are many examples of galaxies where blue-shifted absorption is
detected yet the authors report no significant line emission.  This
includes \ion{Na}{1} flows \citep{rvs05a,martin05,smn+09}, 
\ion{Mg}{2} and \ion{Fe}{2} absorption in
ULIRGS \citep{mb09}, and the extreme \ion{Mg}{2} outflows identified
by \cite{tmd07}.  Similarly, there have been no reports of
resonance line-emission from low-ion transitions in the
LBGs\footnote{\citet{psa+00}
  reported and analyzed \ion{C}{4} line-emission from the lensed LBG
  cb58, attributing this flux to stellar winds.  We caution, however,
  that the galactic-scale wind proposed to exist for cb58 should also
  have contributed to the \ion{C}{4} line-emission.}, and many
$z \lesssim 1$ galaxies show no detectable \ion{Mg}{2} or \ion{Fe}{2}*
emission despite significant blue-shifted absorption (Rubin et al., in prep). 
These non-detections appear to contradict a primary conclusion of 
this paper.   We are motivated, therefore, to
reassess several of the effects that can reduce the line-emission
and consider whether these prevent its detection in many
star-forming galaxies.

Dust is frequently invoked to explain the suppression of line-emission
for resonantly trapped transitions (e.g.\ \lya).  Indeed, a photon
that is trapped for many scatterings within a dusty medium will 
be preferentially extincted relative to a non-resonant photon.  In
$\S$~\ref{sec:dust}, we examined the effects of dust extinction on 
\ion{Mg}{2} and \ion{Fe}{2} emission.  The general result
(Figure~\ref{fig:dust_tau}) is a modest
reduction in flux that scales as (1+\taud)$^{-1}$ instead of
exp(--\taud). Although the \ion{Mg}{2} photons are resonantly
trapped, they require only one to a few scatterings
to escape the wind which limits the effects of
dust.  This reflects the moderate opacity of the \ion{Mg}{2}
doublet (e.g.\ relative to \lya) and also the velocity law of the
fiducial wind model.
In scenarios where the \ion{Mg}{2} photons are
more effectively trapped, dust does suppress
the emission (e.g.\ the ISM+wind model from $\S$~\ref{sec:ISM}).
In contrast to \ion{Mg}{2}, 
the \ion{Fe}{2} resonance photons may be
converted to non-resonant photons that freely escape the wind.
Therefore, a wind model that traps \ion{Mg}{2}
photons for many scatterings does not similarly trap 
\feiid\ photons.  The effects of dust are much 
reduced and, by inference, the same holds for any other set of
transitions that are coupled to a fine-structure level (e.g.\
\ion{Si}{2}/\ion{Si}{2}*).  In summary,
dust does reduce the line flux relative to the continuum, but 
it generally has only a modest effect on the predicted \ion{Mg}{2}
emission and a relatively minor effect on \feiis\ emission.

Another factor that may reduce line-emission is a
non-isotropic wind model. The flux is lower, for example, if one
eliminates the backside to the wind (e.g.\ the source itself could
shadow a significant fraction of the backside).  
For our fiducial model, many of the photons scattered into our
sightline arise nearly normal to the sightline such that a biconicial
wind has significantly lower line-emission
($\S$~\ref{sec:anisotropic}). 
For the anisotropic winds explored in this paper,
the line-emission scales 
roughly as $\Omega/4\pi$ where $\Omega$ is the angular extent of the
wind\footnote{There is also a first order dependence on the actual
  orientation of the wind, with maximal emission when the backside is
  fully viewed.}.  Because we require that the wind 
points toward us, it is difficult to reduce $\Omega$ much
below $2 \pi$ and, therefore, anisotropy reduces the
emission by a factor of order unity.  
Similarly, an anisotropically emitting source would modify the predicted
line-emission.  For example, if the backside of the galaxy were
brighter/fainter then would one predict brighter/fainter emission
relative to the observed continuum.
To reduce the emission, the brightest regions
of the galaxy would need to be oriented toward Earth.  Although this
would not be a
generic orientation, most spectroscopic samples are magnitude-limited
and biased towards detecting galaxies when observed from their
brightest viewpoint.  In principle, this might imply a further
reduction of order unity.
{\bf [Put comments on biconical here]}

Together, dust and anisotropic models may reduce the line-emission
of \ion{Mg}{2} and \ion{Fe}{2}
by one to a few factors of order unity.  For some galaxies, these
effects may explain the absence of significant line-emission,
but these would generally
not be sufficient to preclude detection.

There is another (more subtle) effect that could
greatly reduce the observed line-emission: slit-loss.  As
described in Figure~\ref{fig:fiducial_ifu_mgii}, 
emission from the wind is spatially extended with
a non-zero surface brightness predicted at large radii. This implies 
a non-negligible luminosity emitted beyond the angular extent of the galaxy.  
The majority of observations of star-forming galaxies to date have
been taken through a spectroscopic slit designed to cover
the brightest continuum regions. Slits with angular
extents of $1-1.5''$ subtend roughly $8-15$\,kpc for galaxies at $z>0.5$.
Therefore, a $1''$ slit covering a galaxy with our fiducial wind would cover
less than half the wind.  The result is reduced line-emission
when compared to the galaxy continuum.

We explore the effects of slit-loss as follows.
We model the slit as a perfect, infinitely-long rectangle centered on the
galaxy.  We then tabulate the equivalent width of the line-emission
through slits with a range of widths, parameterized by \rtt: twice the radius
where the wind has a Sobolev optical depth of $\tau_S = 0.2$. 
The line-emission will be weak beyond
this radius because the photons have only a low probability of
being scattered.  For the fiducial wind model, 
the \mgiia\ transition has $\mrtt \approx 15$\,kpc 
(Figure~\ref{fig:fiducial_nvt}).  
The predicted equivalent width in emission \ewe\ relative to the absorption
equivalent width \ewabs\ is presented in
Figure~\ref{fig:obs_slit} for a series of transitions for the fiducial
wind model (black curves).
{\bf [Add radiative model]}
The \ewe/\ewabs\ curves rise very steeply with increasing slit width and then
plateau when the slit width reaches $\approx \mrtt$.  
For the fiducial model, the emission is concentrated toward the source; this
derives from the density and velocity profiles but is also a simple
consequence of geometric projection.
Figure~\ref{fig:obs_slit} reveals similar results for other wind
models. The curves rise so steeply that 
the effects of slit-loss are minor (order unity)
unless the slit-width is very small.   
Nevertheless, the results do motivate
extended aperture observations, e.g.\ integral field unit (IFU)
instrumentation, that would map the wind both spatially and spectrally.

@@

Figure~\ref{fig:obs_ew} demonstrates one observational
consequence: reduced measurements for the absorption
equivalent width \ewabs\ of the flow.  In the case of \ion{Mg}{2},
which has the most effected transitions, \ewabs\ is reduced by
$30-50\%$ from the intrinsic equivalent width.  In turn, one may derive 
a systematically lower optical depth or velocity extent for the wind,
and therefore a lower total mass and kinetic energy.  
The effects are most pronounced for wind
scenarios where the peak optical depth occurs near $\mdv = 0 \mkms$.
Geometric projection limits the majority of scattered photon emission to
have $|\mdv| < 200 \mkms$; therefore,  the absorption profiles
are filled-in primarily at these velocities.

@@

These effects for \ion{Mg}{2} absorption are reduced
for the \ion{Fe}{2} absorption profiles, especially for \feiia.  This
is because a significant fraction (even a majority) of the
\ion{Fe}{2} emission is florescent \ion{Fe}{2}* emission at longer
wavelengths which does not affect the absorption profiles. 
Therefore, the \ion{Fe}{2} absorption equivalent widths (\ewabs) more
closely follow the intrinsic values, one derives more accurate peak
optical depths, and the absorption kinematics more faithfully reflect
the motions of the flow.  Regarding the last point, one also
predicts a velocity offset between the \ion{Fe}{2} and \ion{Mg}{2}
absorption-line centroids (Figure~\ref{fig:obs_vtau}).
%because scattered photons
%preferentially fill-in
%absorption at $v \sim 0 \mkms$ for the \ion{Mg}{2} transitions.  
This affects the analysis of gas related to the ISM of the galaxy and
also material infalling at modest speeds.  
Figure~\ref{fig:obs_vtau} also emphasizes that
the \ion{Mg}{2} profiles may misrepresent the kinematics of the bulk of
the gas.  Analysis of these lines, without consideration of
line-emission, would lead to incorrect conclusions on the energetics
and mass flux of the wind.

@@

The most obvious measures provided by the line-emission are
estimates
of the size and morphology of the outflow \citep{rubin+10c}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths.  The principle
challenge is to achieve sufficient sensitivity to detect the predicted
low
surface brightness emission. As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds that at the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. 
For the fiducial wind model, for example, the azimuthally integrated
flux falls by only a factor of 10 for projected radii of 1 to 15\,kpc.
Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for this analysis because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  The \ion{Mg}{2} lines (especially \mgiib) frequently has the
highest peak and integrated fluxes (Figure~\ref{fig:obs_sb}).  
On the other hand, the \ion{Mg}{2} transitions are more susceptible to
the effects of dust extinction and for some galaxies \feiis\ emission
could be dominant.  In either case, the study of spatially-extended
line-emission from photons scattred by the galactic-scale wind offers
an direct means to study the morphology and radial extent of these
phenomena.  The next generation of large-format optical and infrared
integral-field-units are well-suited to this scientific endeavor
(e.g.\ KCWI on Keck, MUSE on the VLT).

The kinematic characteristics of the line-emission also offer insight into physical
characteristics of the wind.
In $\S$~\ref{sec:dust}, we emphasized that dust extinction
preferentially suppresses photons scattered off the backside of the
wind (e.g.\ $\mdv > 0 \mkms$) so that the line-centroid is shifted to
negative velocities.  Similarly, anisotropic models with reduced
emission off the backside yield emission lines that are centered
blueward of the galaxy's systemic redshift.  These effects are most
prominent in the \feiis\ emission.  The centroids of the \ion{Mg}{2}
lines, meanwhile, are sensitive to the optical depth
profile of the wind.   For example, a wind with flows exceeding
$\approx 600 \mkms$ will shift the \mgiia\ centroid to bluer
wavelengths (e.g.\ Figure~\ref{fig:norm}).
Although these effects are modest (tens of
\kms), they may be resolved by moderate resolution spectroscopy.

The emission-line velocity widths also reveal
characteristics of the wind.  In particular, the \feiis\ emission-line 
widths are
sensitive to the optical depth of the wind at large velocities from
systemic.  This is illustrated in Figure~\ref{fig:obs_edelv} which plots
the 90\%\ velocity width $\Delta v_e$ of \feiic\ versus the velocity
where the wind optical depth is maximal ($v_\tau$).  We find that
$\Delta v_e$ increases with $v_\tau$ such that a large width for
\feiis\ emission requires a wind profile with large optical depths at
$\mdv < -200 \mkms$.  The broadening of \feiis\ emission occurs
because this emission is dominated by single scatterings 
which trace all components of the wind.  {\bf [Also the optical depth profile of
  FeII]}

Lastly, the line flux ratios of pairs of transitions 
are sensitive to characteristics of the wind.
The most obvious example is the relative emisison of the \mgiid\
doublet.  Specifically, an outflow whose velocity exceeds the doublet
spacing ($\Delta v \approx 770 \mkms$) will convert \mgiia\ photons
into \mgiib\ emission resulting in a flux ratio that is inverted
relative to the intrinsic optical depth profiles (e.g.\
Figure~\ref{fig:norm}).  Another important example is the relative
flux of \feiis\ emission relative to \ion{Mg}{2}.  
A dusty medium with velocity near systemic (e.g.\ an
ISM component) may significantly suppress \ion{Mg}{2} emission yet
have a significant \feiis\ line flux because the latter is not
resonantly trapped.

@@

In this paper, we have explored the predicted absorption and
emission-line profiles of a set of simple galactic-scale outflow
models.  Our analysis focused on the \mgiid\ and \ion{Fe}{2} UV1
multiple of rest-frame UV transitions, but the results translate to
most other transitions used to probe cool gas.  Our primary
findings are summarized as follows:

@@

\item Isotropic, dust-free wind models conserve photon flux.
  Therefore, the blue-shifted absorption-line profiles commonly
  observed in star-forming galaxies are predicted to be accompanied by
  emission-lines with similar equivalent width.  This is even true for
  anisotropic and dust-extincted scenarios.

\item The line-emission occurs preferentially at the systemic velocity
  of the galaxy and `fills-in' the absorption profiles at velocity
  offsets $|\mdv| < 200 \mkms$ from systemic.  For transitions that
  are only coupled to the ground-state (e.g.\ \ion{Mg}{2}, \lya,
  \ion{Na}{1}). The observed absorption profiles may be significantly
  different from the intrinsic optical depth profile.

@@

\item Resonance transitions that are strongly coupled to
non-resonant lines (e.g.\ \ion{Fe}{2}, \ion{Si}{2}) offer the best
characterization of the opacity of the wind and also material 
with velocities near systemic.  The emission will generally be dominated by the
optically-thin, fine-structure transitions.

\item  Dust extinction has only modest effects 
  in models which predict only a few scatterings.  Models
  with high opacity at small radii and at systemic velocity (e.g.\ an
  optically thick, ISM component)  can effectively extinguish
  resonantly trapped emission (\ion{Mg}{2}) but have more modest
  effect on non-resonant lines (\feiis).

\item We examined two scenarios designed to mimic the wind model
  proposed by S10 for $z \sim 3$ Lyman break galaxies.  These models
  predict substantial line-emission from scattered photons that
  greatly modify the predicted line-profiles.  
  These models do not reproduce the observed line-profiles.

%\item In fact, the
%\feiia\ profile very nearly matches the profile without re-emission 
%(compare to the dotted lines); this transition provides a
%very good description of the intrinsic ISM+wind optical depth profile.  
%We conclude that resonant transitions that are coupled to (multiple)
%non-resonant, electric dipole transitions offer the best
%diagnosis of ISM absorption.
